---
title: 'Using Data to Improve a Marketing Promotion'
author: "Matt Jackson"
date: "December 10, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries

```{r warning=FALSE, message=FALSE}
library("ggplot2")
library("Hmisc")
library("dplyr")
library("lubridate")
```

## Load Transformed Dodgers Dataset

```{r}
# Load Dataset
dodgers <- read.csv("../dodgers_transformed.csv")

# Convert Variables to the Right Types
dodgers$date <- as.Date(dodgers$date)
dodgers$month <- month(dodgers$date)
dodgers$cap <- as.logical(dodgers$cap)
dodgers$shirt <- as.logical(dodgers$shirt)
dodgers$fireworks <- as.logical(dodgers$fireworks)
dodgers$bobblehead <- as.logical(dodgers$bobblehead)

# Reorder Days of the Week
dodgers$day_of_week <- factor(dodgers$day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))

# Create Promo Column
dodgers$promo <- FALSE
dodgers$promo[dodgers$cap==TRUE] <- TRUE
dodgers$promo[dodgers$shirt==TRUE] <- TRUE
dodgers$promo[dodgers$fireworks==TRUE] <- TRUE
dodgers$promo[dodgers$bobblehead==TRUE] <- TRUE
```

## Plot Scatterplot for Attendance over Time with Days of the Week Colored

```{r}
ggplot(dodgers, aes(x=date, y=attend, color=day_of_week)) +
  geom_point(size=2) +
  xlab("Date") +
  ylab("Attendance (Number of People)") + 
  labs(color="Day of the Week")
```

## Plot Boxplots for Each Day of the Week

```{r}
ggplot(dodgers, aes(x=day_of_week, y=attend)) +
  geom_boxplot() +
  xlab("Day of the Week") +
  ylab("Attendance (Number of People)")
```

## Plot of Promos Being Ran on Mondays

```{r}
dogers_monday <- subset(dodgers, day_of_week=="Monday")

ggplot(dogers_monday, aes(x=promo)) +
  geom_bar() +
  xlab("Promotion Ran?") +
  ylab("Count")
```

## Simple Regression Model with Just Day of the Week

```{r}
simple <- lm(attend ~ day_of_week, data=dodgers)
summary(simple)
```

## Create DF to Predict New Attendance Values

```{r}
days <- data.frame("day_of_week" = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
days$day_of_week <- factor(days$day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```

## Predict Values with Simple Regression Model

```{r}
predict(simple, days)
```

## Multiple Regression Model with Day of the Week and Promo

```{r}
multiple <- lm(attend ~ day_of_week + promo, data=dodgers)
summary(multiple)
```

## Add Column for Promo (No Promos Being Run)

```{r}
days$promo <- c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)
```

## Predict Values with Multiple Regression Model (No Promos)

```{r}
days_with_no_promos <- predict(multiple, days)
```

## Add Column for Promo (Promos Being Run)

```{r}
days$promo <- c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE)
```

## Predict Values with Multiple Regression Model (Promos)

```{r}
days_with_promos <- predict(multiple, days)
```

## Calculate Difference 

```{r}
days_with_promos - days_with_no_promos
```

## Conclusion
When the scatterplot and boxplots were produced, it is clear that Monday has the lowest attendance. After applying a simple and multiple linear regression model, that still reamined true somewhat (in the mutliple model Sunday had slightly lower attendance). Running a multiple linear regression model taking into account how promos affect attendance, on average the attendance lift is about 10,600 people.